---
title: "Machine Learning Course Project"
author: "Dean Wise"
date: "April 13, 2016"
output: html_document
---

##Prepping/Cleaning the Data

First I am going to load libraries, set the seed,  and load the data:

```{r}
library(caret); library(kernlab)
set.seed(83)
training <- read.csv("pml-training.csv", na.strings=c("", "NA", "#DIV/0!"))
test <- read.csv("pml-testing.csv", na.strings=c("", "NA", "#DIV/0!"))
```

Next let's partition the data:

```{r}
inTrain <- createDataPartition(training$classe, p=0.75, list=FALSE)
training2 <- training[inTrain, ]
testing2 <- training[-inTrain, ]
dim(training2)
dim(testing2)
```

We can now use nearZeroVar() to get rid of near zero variance variables, which can often break models we would want to use. 

```{r}

nzv <- nearZeroVar(training2, saveMetrics=TRUE)
training2 <- training2[,nzv$nzv==FALSE]

nzv<- nearZeroVar(testing2,saveMetrics=TRUE)
testing2 <- testing2[,nzv$nzv==FALSE]

```

There is now an index column in the training and testing data, which we need to remove:
```{r}
training2 <- training2[c(-1)]
testing2 <- testing2[c(-1)]
```

There are a lot of columns that still have a large portion of NA's. We are going to remove any column that is made up of at least 75% NA's. Once we have those columns, we can them map them onto the test set so the columns align. 

Remove columns with 75% NA's:
```{r}

training3 <- training2
for(i in 1:length(training2)) {
    if( sum( is.na( training2[, i] ) ) /nrow(training2) >= .75) {
        for(j in 1:length(training3)) {
            if( length( grep(names(training2[i]), names(training3)[j]) ) == 1)  {
                training3 <- training3[ , -j]
            }   
        } 
    }
}

training2 <- training3

```

Now we want to map the columns from the training2 set onto the testing data sets:
```{r}

headers <- colnames(training2)
testing2 <- testing2[headers]
test <- test[headers[1:57]]

```

Now the only issue left is that the classes of the columns between our training and testing do not align. So we need to make them the same class:

```{r}

for (i in 1:length(test) ) {
    for(j in 1:length(training2)) {
        if( length(grep(names(training2[i]), names(test)[j])) == 1)  {
            class(test[j]) <- class(training2[i])
        }      
    }      
}

test <- rbind(training2[2, -58] , test)
test <- test[-1,]

```


##Predicting With the Data


#https://rpubs.com/DIM302/PracticalMachineLearning























































































